### 2008-08-18 doc/a-rex/arex_tech_doc.tex
A-REX tech doc update. Language fixes from Steffen and spellchecker. Citations if available.

A-REX is one of ARC middleware componets implementing functions of
so called \emph{Computing Element} (CE).

The aim of the A-REX is to take care of job pre- and post-processing.
It provides functionality to stage-in files containing input data
and program modules from wide range of sources and transfer or store
output results.

The A-REX implements Web Service (WS) interface which provides a way
to submit and control computational tasks (jobs) to be executed by
the A-REX and undelying Local Resource Management System.

\textbf{You should use this document for advanced configuration purposes
and understanding of internals of the aforementioned tools. For general
instalaltion and configuration of whole system refer to other documents
avaialble at \url{http://www.nordugrid.org/papers.html}. }

@@

A job is a set of input files (which may include executables), a main
executable and a set of output files. The process of gathering input
files, executing a job, and transferring/storing output files is called
a \emph{session}.

Each job gets a directory on the CE called the \emph{session directory}
(SD). Input files are gathered in the SD. The job is supposed to produce
new data files in the SD as well. The A-REX does not guarantee the
availability of any other places accessible by the job other than
SD (unless such a place is part of a requested Runtime Environement).
The SD is also the only place which is controlled by the A-REX. It
is accessible by the user from outside through HTTP(S) protocol. Any
file created outside the SD is not controlled by the A-REX. Any exchange
of data between client and A-REX (including also program modules)
is done through HTTP(S) protocol. A URL for accessing input/output
files is  obtained through the WS Local Information Description Interface
(LIDI) of A-REX.

Each job gets an identifier (\textit{jobid}). This is a handle - WS-Addressing
\cite{jsdl} XML document - which identifies the job in the A-REX
and the Information Interface.

Jobs are initiated and controlled through WS interface. Complete job
descriptions (JD)are passed to the A-REX through WS in JSDL\cite{jsdl}
coded description. Input data files and job executables are transferred
separately through the same interface, as described in the next section.

@@

\begin{itemize}
\item Download is initiated by the A-REX -- This is the case for files defined
in the JD (with name and source). The A-REX alone is responsible to
make sure that all required files will be available in the SD.\\
The supported protocols for sources at the moment are (in case of
full installation): GridFTP, FTP, HTTP, HTTPS (HTTP over SSLv3). Also
some less standard sources are supported. Those are described below.
The A-REX fully relies on HED framework \cite{hed} for data transfering
capabilities. Hence actual set of supported protocols depends on installed
Data Management Components of the HED.
\item Upload is initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the A-REX. Examples of such
files are the main executable of the job, the job's input files, etc.
These files can (and should) also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE based on the A-REX. Access to AFS, NFS, FTP, HTTP and
any other remote data transport during execution of a job is not guaranteed
(at least not by A-REX).

@@

\begin{itemize}
\item Files which are supposed to be moved to a \emph{Storage Element} (SE)
and optionally registered in some \emph{Indexing Service} like the
Globus \emph{Replica Location Service} (RLS) -- The A-REX takes care
of these files. They have to be specified in the JD. If the job fails
during any stage of processing, no attept is made to transfer those
files to their final destination, unless the option \emph{preserve=yes}
is specified in their URLs.
\item Files which are supposed to be fetched by the user -- The user has
to use a tool like the UI to obtain these files. They \textbf{must}
also be specified in the JD.
\end{itemize}
All files not specified in the JD are deleted after job execution
finished. If job execution fails for any reason (if exit code of main
executable is not 0) all files from furst group are transfered to
second one.

@@

Configuration can put limits on the amout of simultaneous jobs in
some states. If such a limit is reached, a job ready to enter into
the state in question will stay in it's current state waiting for
a free slot. This situation is presented by additional state mark
\textbf{PENDING} to the current state name in the job's status desription.

@@

\begin{itemize}
\item \textbf{Accepted} - In this state the job has been submitted to a
CE but is not processed yet. The A-REX will analyze the JD and move
to the next stage. If the JD can not be processed the job will be
canceled and moved to the state \textbf{Finishing}.
\item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
The A-REX is downloading the files specified in the JD and is waiting
for files which are supposed to be uploaded by the UI. If all files
are successfully gathered the job moves to the next state. If \textbf{any}
file can't be downloaded or it takes the UI too long to upload a file,
the job moves to \textbf{Finishing} state. It is possible to put a
limit on the number of simultaneous \textbf{Preparing} jobs. If this
limit is exceeded, jobs ready to enter the \textbf{Preparing} state
 will stay in the \textbf{Accepted} state, but prefixed with the PENDING:
mark. Exceptions are jobs which have no files to be downloaded. These
are processed out of limits.
\item \textbf{Submiting} - The job is being passed for execution to the
\emph{Local Resource Management System} (LRMS). The correspoding backends
for many LRMSs are provided with the default installation. If the
local job submission is successful the job moves to the \textbf{Executing}
state. Otherwise it moves to \textbf{Finishing}. It is possible to
limit the agregate number of jobs in \textbf{Submiting} and \textbf{Executing}
states.
\item \textbf{Executing} - The job is queued or being executed in the LRMS.
The A-REX takes no actions except waiting until the job finishes.
\item \textbf{Killing} - Necessary action to cancel the job in the LRMS
is being taken.
\item \textbf{Finishing} - The output data is being processed (stage-out).
Specified data files are moved to the specified SEs and are optionally
registered at an Indexing Service. The user can download data files
from the SD by using the UI or other adequate tool. All the files
not specified as output files are removed from the SD at very beginning
of this state. It is possible to limit number of simultaneous jobs
in this state.
\item \textbf{Finished} - No more processing is performed by the A-REX.
The user can continue to download data files from the SD. The SD is
kept available for some time (default is 1 week). After that the job
is moved to the state \textbf{Deleted}. The 'deletion' time can be
obtained by querying the Informatrion Interface of the A-REX. If a
job was moved to \textbf{Finished} because of failure, it may be restarted
on request of a client. When restarted, a job is moved to the state
previous to the one in which it failed and is assigned mark PENDING.
This is needed in order to not break the configuration limits. Exception
is a job failed in \textbf{Executing} state and lacking input files
specified in JD. Such a job is treated like failed in \textbf{Preparing}
state.
\item \textbf{Deleted} - The job is moved to this state if the user have
not requested job to be cleaned before the SD's lifetime expires.
Only minimal subset of information about such job is kept. The SD
is not available anymore.
\end{itemize}
In case of a failure special processing is applied to output files.
All specified output files are treated as \textbf{downloadable by
the user}. No files will be moved to their destination SE.

@@

\begin{itemize}
\item \textit{job.ID.status} - current state of the job. This is a plain
text file containing a single word representing the internal name
of current state of the job. Possible values and corrsponding external
job states are :

@@

\begin{itemize}
\item \textit{subject} - user certificate's subject, also known as the distinguished
name (DN)
\item \textit{starttime} - GMT time when the job was accepted represented
in the Generalized Time format of LDAP 
\item \textit{lifetime} - time period to preserve the SD after the job has
finished in seconds
\item \textit{cleanuptime} - GMT time when the job should be removed from
the cluster and it's SD deleted in Generalized Time format
\item \textit{notify} - email addresses and flags to send mail to about
the job specified status changes
\item \textit{processtime} - GMT time when to start processing the job in
Generalized Time format
\item \textit{exectime} - GMT time when to start job execution in Generalized
Time format
\item \textit{expiretime} - GMT time when the credentials delegated to the
job expire in Generalized Time format
\item \textit{rerun} - number of retries left to rerun the job
\item \textit{jobname} - name of the job as supplied by the user
\item \textit{lrms} - name of the LRMS backend to be used for local submission
\item \textit{queue} - name of the queue to run the job at
\item \textit{localid} - job id in LRMS (appears only after the job has
reached state \textbf{InLRMS})
\item \textit{args} - executable name followed by a list of command-line
arguments
\item \textit{downloads} - number of files to download into the SD before
execution
\item \textit{uploads} - number of files to upload from the SD after execution
\item \textit{gmlog} - directory name which holds files containing information
about the job when accessed through GridFTP interface
\item \textit{clientname} - name (as provided by the user interface) and
ip address:port of the submitting client machine
\item \textit{clientsoftware} - version of software used to submit the job
\item \textit{sessiondir} - the job's SD
\item \textit{failedstate} - state in which job failed (available only if
it is possible to restart the job)
\item \textit{jobreport} - URL of a user requested \emph{logger service}.
The A-REX will also send job records to this service in addition to
the default logger service configured in the configuration (This feature
does not work in current implementation).
\end{itemize}
\end{itemize}
This file is filled partially during job submission and fully when
the job moves from the \textbf{Accepted} to the \textbf{Preparing}
state.

@@

\textit{url} - ordinary URL for gsiftp, ftp, http, https protocols
with the addition of '\textbf{replica location service url}' (RLS
URL).\\
Each URL can contain additional options.

@@

\end{itemize}
This file is used by the '\textbf{\textit{downloader}}' utility. Files
with 'url' will be downloaded to the SD and files with 'file description'
will simply be checked to exist. Each time a new \textbf{valid} file
appears in the SD it is removed from the list and \textit{job.ID.input}
is updated. Any external tool can thus track the process of collecting
input files by checking \textit{job.ID.input.}

\begin{itemize}
\item \textit{job.ID.output} - list of output files. Each line contains
1 or 2 values separated by a space. First value is the name of the
file relative to the SD. The second value, if present, is a URL. Supported
URLs are the same as those supported by job.ID.input.
\end{itemize}
This file is used by the '\textbf{\textit{uploader}}' utility. Files
with \textit{url} will be uploaded to SE and remaining files will
be left in the SD. Each time a file is uploaded it is removed from
the list and \textit{job.ID.output} is updated. Files not mentioned
as output files are removed from the SD at the the beginning of the
\textbf{Finishing} state.

@@

A-REX Web Service Interface provides means to submit a description
of a computational job to a computing resource, to stage-in additional
data, to monitor and control processing of jobs, and obtain data produced
during the execution of a job. The WS Interface is built and deployed
inside Hosting Envirnoment Daemon (HED) infrastructure\cite{hed}.

@@

The job submission and control interface is based on document produced
by the OGF OGSA Basic Execution Services (BES) Working Group\cite{bes}.
The exchange of SOAP messages is performed over the HTTP(S) protocol.
The BES interface is represented by two port-types - BES-Management
and BES-Factory. The former is made to control the A-REX service itself
and thus defines operations to start and stop the functionality of
the BES service. The A-REX does not implement remote control of service
functionality. Hence the BES-Management port-type is not functional.
The BES-Factory port-type provides operations to submit new jobs (to
create an activity in terms of BES) and to monitor its state. It also
has an ability to provide information about the service. A-REX fully
implements the functionality of this port-type.

@@

Since BES allows implementators to extend their initial activity states
with additional sub-states, A-REX defines a set of sub-states of activity
processing in addition to those defined by the BES, as listed in Table
\ref{tab:Job-states-definitions}. Their meaning is described in section
\ref{Section: Job Flow}.

@@

The A-REX also supportd the Delegation Interface (see Appendix D).
This is a common purpose interface to be used by ARC services which
accepts delegated credentials from clients. The Delegation Interface
implements two operations: initiatialization of credentials delegation
(DelegateCredentialsInit) and update/renewal of credentials (UpdateCredentials).

@@

\begin{lyxlist}{00.00.0000}
\item [{\emph{nordugrid}}] - desription of computing resource that uses
NorudGrid LDAP schema {[}] converted to XML document.
\item [{\emph{Domains}}] \emph{-} description of a computation resource
that uses Glue2 schema.
\end{lyxlist}
All information can be accessed either through requests on particular
resources or through XPath queries using WS-Resource Properties operations.

@@

A-REX supports the following elements from the JSDL version 1.0 specification\cite{jsdl}
including POSIX Applications extension and JSDL HPC Profile Application
Extension {[}]:

@@

\begin{lyxlist}{00.00.0000}
\item [{Filename}] - name of the data file on the executing node.
\item [{\emph{Source}}] - source where the file will be taken from before
execution.
\item [{\emph{Target}}] \emph{-} destination the file will be delivered
to after execution.
\end{lyxlist}
\item [{\emph{Input}\ (POSIX,HPC)}] - file to be used as standard input
for the executable.
\item [{\emph{Output}\ (POSIX,HPC)}] - file to be used as standard output
for the executable.
\item [{\emph{Error}\ (POSIX,HPC)}] - file to be used as standard error
for the executable.
\item [{\emph{MemoryLimit}\ (POSIX)}] - amount of physical memory needed
for execution.
\item [{\emph{TotalPhysicalMemory}}] - same as MemoryLimit.
\item [{\emph{IndividualPhysicalMemory}}] - same as MemoryLimit.
\item [{\emph{CPUTimeLimit}\ (POSIX)}] - maximal amount of CPU time needed
for execution.
\item [{\emph{TotalCPUTime}}] - same as CPUTimeLimit.
\item [{\emph{IndividualCPUTime}}] - same as CPUTimeLimit.
\item [{\emph{WallTimeLimit}\ (POSIX)}] - amount of clock time needed
for execution.
\item [{\emph{TotalCPUCount}}] \emph{-} number of CPUs needed for execution.
\item [{\emph{IndividualCPUCount}}] - same as TotalCPUCount.
\item [{\emph{Environment}\ (POSIX)}] - environement variable.
\end{lyxlist}

@@

\begin{itemize}
\item \textit{hardsize} -size of file system for storing cached data
\item \textit{hardfree} - amount of disc space available on that file system
\item \textit{softsize} - if cache exceeds this size files are started being
removed 
\item \textit{softfree} - space left till softsize (can be negative)
\item \textit{claimed} - space used by files claimed by running jobs
\item \textit{unclaimed} - space used by files not being currently used
by any job
\end{itemize}
\item \textit{\#\#\#\#\#\#\#\#.info} - stores state of file (\textit{\#\#\#\#\#\#\#\#}
stands for 8 digits). State is represented by one character:\\
\textit{c} - just created, content is empty.\\
\textit{f} - failed to download (treated same as 'c').\\
\textit{r} - ready to be used, content is valid.\\
\textit{d} - being downloaded. 'd' is followed by identifier of application/job
downloading that file. During content's download this file has write
lock set.
\item \textit{\#\#\#\#\#\#\#\#.claim} - stores list of identifiers of applications/jobs
using this file. Identifiers are stored one per line.
\item \textit{\#\#\#\#\#\#\#\#} - files storing content of corresponding
URL. These can be stored in separate directory.
\end{itemize}
Files \textit{list, old , new} and \textit{\#\#\#\#\#\#\#\#.info}
has to be stored on filesystem which has support for files' locking.

@@

If a job requests an input file which is subject for cacheing, it
is stored in the cache directory instead of the SD. The file is made
available to the job by either soft-linking it in the SD or copying
it to the SD. The latter option is more secure and hence advised.

@@

Before giving access to an already cached file, the A-REX contacts
the original file source to check if the user has suficient access
rights to the file. Not all protocols make this check possible. Also,
file creation or validity times are checked to make sure the cached
file is fresh. If it is impossible to obtain creation and invalidation
times, the file is invalidated 24 hours after it was downloaded.

@@

\begin{itemize}
\item \textit{libarex.so} - The main module providing main functionality
and web interface. It is implemented as HTTP and SOAP service inside
HED. It is responsible for processing jobs, moving them through states
and running other modules.
\item \textit{downloader} - This is a module responsible for gathering input
files in the SD. It processes the \textit{job.ID.input} file and updates
it.
\item \textit{uploader} - This module is responsible for delivering output
files to the specified SEs and registration at an Indexding Service
(like RLS) as needed. It processes and updates the \textit{job.ID.output}
file.
\item \emph{cache-register} - Utility to register cached data into an Indexing
Service. It reads and modifies cache informational files \emph{old}
and \emph{new} (as described in section \ref{sec:cache}). Configuration
is read directly from the A-REX's configuration file (see section
\ref{SubSection: ConfigFile}). It is run by the A-REX every 5 minutes.
\item \emph{frontend-info-collector} - Utility to gather information about
the frontend. It puts collected information into the \emph{job.ID.diag}
file.
\item \emph{gm-kick} - Sends a signal to the A-REX though a FIFO file to
wake it up. It's used to increase responsiveness of A-REX.
\item \emph{cluster+qju.pl} - Collects and generates information about computing
resource as XML document in NoruGrid and Glue 2 format.
\end{itemize}
The following modules are always run under the Unix account to which
a Grid user is mapped.

\begin{itemize}
\item \textit{smtp-send.sh} and \textit{smtp-send} - These are the modules
responsible for sending e-mail notifications to the user. The format
of the mail messages can be easily changed by editing the simple shell
script \textit{smtp-send.sh}. 
\item \textit{submit-{*}-job} - Here {*} stands for the name of the LRMS.
Curently supported LRMS are PBS/Torque, Condor and SGE. Also \emph{fork}
pseudo-LRMS is supported for testing purposes. This module is responsible
for job submission to the LRMS.
\item \textit{cancel-{*}-job} - This script is for canceling jobs which
have been already submitted to the LRMS.
\item \textit{scan-{*}-job} -This shell script is responsible for notifying
the A-REX about completion of jobs. It's implementation for PBS uses
server logs to extract information about jobs. If logs are not available
it uses the less reliable \emph{qstat} command for that. Other backends
use different techniques.
\end{itemize}
In addition, there is also an administration utility:

@@

Due to historical resons configuration of the A-REX is split into
2 parts. The small part is located inside HED configuration (see appendix
B) for schema and description of supported elements. It refers to
file containing main part of configuration. The default location of
main configuration file is \textit{/etc/arc.conf .}

@@

requires installation of A-REX and all realted software to be accessible

@@

\textit{max\_transfered\_files} - maximal number of files being transfered

@@

are registred. Registration is done to same service used for obtaining
information about file. For this opeartion credentials of the A-REX

@@

\textit{log} - write information about result into logfile and continue

@@

\textit{fail} - write information about result into logfile and cancel

@@

a user to which Grid identity is mapped in order to access filesystem

@@

max\_inactivity} - specifies how long/slow data fransfer is allowed

@@

data is receved for longer than \textit{max\_inactivity} seconds.

@@

Please read \emph{User Manual {[}]} for usage instructions.

>>>>

A-REX is one of ARC middleware components that implements functions
of so called \emph{Computing Element} (CE).

It takes care of job pre- and post-processing. This comprises the
stage-in of files containing input data or program modules from wide
range of sources and the transfer or storing of output results.

The A-REX implements a Web Service (WS) interface which provides a
way to submit and control computational tasks (jobs) to be executed
by the A-REX and he underlying Local Resource Management System.

\textbf{You should use this document for advanced configuration purposes
and understanding of internals of the aforementioned tools. For general
installation and configuration of whole system refer to other documents
available at \url{http://www.nordugrid.org/papers.html}. }

@@

On the computing element a job is described as a set of input files
(which may include executables), a main executable and a set of output
files. The process of gathering input files, executing a job, and
transferring/storing output files is called a \emph{session}.

Each job gets a directory on the CE called the \emph{session directory}
(SD). Input files are gathered in the SD. The job is supposed to produce
new data files in the SD as well. The A-REX does not guarantee the
availability of any other places accessible by the job other than
SD (unless such a place is part of a requested Runtime Environment).
The SD is also the only place which is controlled by the A-REX. It
is accessible by the user from outside through the HTTP(S) protocol.
Any file created outside the SD is not controlled by the A-REX. Any
exchange of data between client and A-REX (including also program
modules) is performed via HTTP(S). A URL for accessing input/output
files is  obtained through the WS Local Information Description Interface
(LIDI) of A-REX.

Each job gets an identifier (\textit{jobid}). This is a handle - WS-Addressing
\cite{ws-addr-soap} XML document - which identifies the job in the
A-REX and the Information Interface.

Jobs are initiated and controlled through the WS interface. Complete
job descriptions (JD)are passed to the A-REX through WS in JSDL\cite{jsdl}
coded description. Input data files and job executables are transferred
separately through the same interface, as described in the next section.

@@

\begin{itemize}
\item Download is initiated by the A-REX -- This is the case for files defined
in the JD (with name and source). The A-REX alone is responsible to
ensure that all required files will be available in the SD.\\
The supported protocols for sources at the moment are (in case of
full installation): GridFTP, FTP, HTTP, HTTPS (HTTP over SSLv3). Also
some less standard sources are supported. Those are described below.
The A-REX fully relies on HED framework \cite{hed} for data transferring
capabilities. Hence actual set of supported protocols depends on installed
Data Management Components of the HED.
\item Upload is initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the A-REX. Examples of such
files are the main executable of the job, the job's input files, etc.
These files can (and should) also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE based on the A-REX. Access to AFS, NFS, FTP, HTTP and
any other remote data transport during execution of a job is not guaranteed
(at least not by A-REX).

@@

\begin{itemize}
\item Files which are supposed to be moved to a \emph{Storage Element} (SE)
and optionally registered in some \emph{Indexing Service} like the
Globus \emph{Replica Location Service} (RLS) -- The A-REX takes care
of these files. They have to be specified in the JD. If the job fails
during any stage of processing, no attempt is made to transfer those
files to their final destination, unless the option \emph{preserve=yes}
is specified in their URLs.
\item Files which are supposed to be fetched by the user -- The user has
to use a tool like the UI to obtain these files. They \textbf{must}
also be specified in the JD.
\end{itemize}
All files not specified in the JD are deleted after job execution
finished. If job execution fails for any reason (if exit code of main
executable is not 0) all files from first group are transferred to
second one.

@@

Configuration can put limits on the amount of simultaneous jobs in
some states. If such a limit is reached, a job ready to enter into
the state in question will stay in it's current state waiting for
a free slot. This situation is presented by additional state mark
\textbf{PENDING} to the current state name in the job's status description.

@@

\begin{itemize}
\item \textbf{Accepted} - In this state the job has been submitted to a
CE but is not processed yet. The A-REX will analyze the JD and move
to the next stage. If the JD can not be processed the job will be
canceled and moved to the state \textbf{Finishing}.
\item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
The A-REX is downloading the files specified in the JD and is waiting
for files which are supposed to be uploaded by the UI. If all files
are successfully gathered the job moves to the next state. If \textbf{any}
file can't be downloaded or it takes the UI too long to upload a file,
the job moves to \textbf{Finishing} state. It is possible to put a
limit on the number of simultaneous \textbf{Preparing} jobs. If this
limit is exceeded, jobs ready to enter the \textbf{Preparing} state
 will stay in the \textbf{Accepted} state, but prefixed with the PENDING:
mark. Exceptions are jobs which have no files to be downloaded. These
are processed out of limits.
\item \textbf{Submiting} - The job is being passed for execution to the
\emph{Local Resource Management System} (LRMS). The corresponding
backends for many LRMSs are provided with the default installation.
If the local job submission is successful the job moves to the \textbf{Executing}
state. Otherwise it moves to \textbf{Finishing}. It is possible to
limit the aggregate number of jobs in \textbf{Submiting} and \textbf{Executing}
states.
\item \textbf{Executing} - The job is queued or being executed in the LRMS.
The A-REX takes no actions except waiting until the job finishes.
\item \textbf{Killing} - Necessary action to cancel the job in the LRMS
is being taken.
\item \textbf{Finishing} - The output data is being processed (stage-out).
Specified data files are moved to the specified SEs and are optionally
registered at an Indexing Service. The user can download data files
from the SD by using the UI or other adequate tool. All the files
not specified as output files are removed from the SD at very beginning
of this state. It is possible to limit number of simultaneous jobs
in this state.
\item \textbf{Finished} - No more processing is performed by the A-REX.
The user can continue to download data files from the SD. The SD is
kept available for some time (default is 1 week). After that the job
is moved to the state \textbf{Deleted}. The 'deletion' time can be
obtained by querying the Information Interface of the A-REX. If a
job was moved to \textbf{Finished} because of failure, it may be restarted
on request of a client. When restarted, a job is moved to the state
previous to the one in which it failed and is assigned mark PENDING.
This is needed in order to not break the configuration limits. Exception
is a job failed in \textbf{Executing} state and lacking input files
specified in JD. Such a job is treated like failed in \textbf{Preparing}
state.
\item \textbf{Deleted} - The job is moved to this state if the user have
not requested job to be cleaned before the SD's lifetime expires.
Only minimal subset of information about such job is kept. The SD
is not available anymore.
\end{itemize}
In case of a failure special processing is applied to output files.
All specified output files are treated as \textbf{downloadable by
the user}. No files will be moved to their destination SE.

@@

\begin{itemize}
\item \textit{job.ID.status} - current state of the job. This is a plain
text file containing a single word representing the internal name
of current state of the job. Possible values and corresponding external
job states are :

@@

\begin{itemize}
\item \textit{subject} - user certificate's subject, also known as the distinguished
name (DN)
\item \textit{starttime} - GMT time when the job was accepted represented
in the Generalized Time format of LDAP 
\item \textit{lifetime} - time period to preserve the SD after the job has
finished in seconds
\item \textit{cleanuptime} - GMT time when the job should be removed from
the cluster and it's SD deleted in Generalized Time format
\item \textit{notify} - email addresses and flags to send mail to about
the job specified status changes
\item \textit{processtime} - GMT time when to start processing the job in
Generalized Time format
\item \textit{exectime} - GMT time when to start job execution in Generalized
Time format
\item \textit{expiretime} - GMT time when the credentials delegated to the
job expire in Generalized Time format
\item \textit{rerun} - number of retries left to rerun the job
\item \textit{jobname} - name of the job as supplied by the user
\item \textit{lrms} - name of the LRMS backend to be used for local submission
\item \textit{queue} - name of the queue to run the job at
\item \textit{localid} - job id in LRMS (appears only after the job has
reached state \textbf{InLRMS})
\item \textit{args} - executable name followed by a list of command-line
arguments
\item \textit{downloads} - number of files to download into the SD before
execution
\item \textit{uploads} - number of files to upload from the SD after execution
\item \textit{gmlog} - directory name which holds files containing information
about the job when accessed through GridFTP interface
\item \textit{clientname} - name (as provided by the user interface) and
IP address:port of the submitting client machine
\item \textit{clientsoftware} - version of software used to submit the job
\item \textit{sessiondir} - the job's SD
\item \textit{failedstate} - state in which job failed (available only if
it is possible to restart the job)
\item \textit{jobreport} - URL of a user requested \emph{logger service}.
The A-REX will also send job records to this service in addition to
the default logger service configured in the configuration (This feature
does not work in current implementation).
\end{itemize}
\end{itemize}
This file is filled partially during job submission and fully when
the job moves from the \textbf{Accepted} to the \textbf{Preparing}
state.

@@

\emph{URL} represents location from which file can be downloaded.
Each URL can contain additional options.

@@

\end{itemize}
This file is used by the '\textbf{\textit{downloader}}' utility. Files
with \emph{URL} will be downloaded to the SD and files with 'file
description' will simply be checked to exist. Each time a new \textbf{valid}
file appears in the SD it is removed from the list and \textit{job.ID.input}
is updated. Any external tool can thus track the process of collecting
input files by checking \textit{job.ID.input.}

\begin{itemize}
\item \textit{job.ID.output} - list of output files. Each line contains
1 or 2 values separated by a space. First value is the name of the
file relative to the SD. The second value, if present, is a URL. Supported
URLs are the same as those supported by job.ID.input.
\end{itemize}
This file is used by the '\textbf{\textit{uploader}}' utility. Files
with \textit{URL} will be uploaded to SE and remaining files will
be left in the SD. Each time a file is uploaded it is removed from
the list and \textit{job.ID.output} is updated. Files not mentioned
as output files are removed from the SD at the the beginning of the
\textbf{Finishing} state.

@@

A-REX Web Service Interface provides means to submit a description
of a computational job to a computing resource, to stage-in additional
data, to monitor and control processing of jobs, and obtain data produced
during the execution of a job. The WS Interface is built and deployed
inside the Hosting Environment Daemon (HED) infrastructure\cite{hed}.

@@

The job submission and control interface is based on a document produced
by the OGF OGSA Basic Execution Services (BES) Working Group\cite{ogsa-bes}.

The exchange of SOAP messages is performed via HTTP(S). The BES interface
is represented by two port-types - BES-Management and BES-Factory.
The former is made to control the A-REX service itself and thus defines
operations to start and stop the functionality of the BES service.
The A-REX does not implement remote control of service functionality.
Hence the BES-Management port-type is not functional. The BES-Factory
port-type provides operations to submit new jobs (to create an activity
in terms of BES) and to monitor its state. It also has an ability
to provide information about the service. A-REX fully implements the
functionality of this port-type.

@@

Since BES allows implementations to extend their initial activity
states with additional sub-states, A-REX defines a set of sub-states
of activity processing in addition to those defined by the BES, as
listed in Table \ref{tab:Job-states-definitions}. Their meaning is
described in section \ref{Section: Job Flow}.

@@

The A-REX also supports the Delegation Interface (see Appendix D).
This is a common purpose interface to be used by ARC services which
accepts delegated credentials from clients. The Delegation Interface
implements two operations: initialization of credentials delegation
(DelegateCredentialsInit) and update/renewal of credentials (UpdateCredentials).

@@

\begin{lyxlist}{00.00.0000}
\item [{\emph{nordugrid}}] - description of computing resource that uses
NorudGrid LDAP schema \cite{is} converted to XML document.
\item [{\emph{Domains}}] \emph{-} description of a computation resource
that uses Glue2 schema.
\end{lyxlist}
All information can be accessed either through requests on particular
resources or through XPath queries using WS-Resource Properties operations.

@@

A-REX supports the following elements from the JSDL version 1.0 specification\cite{jsdl}
including POSIX Applications extension and JSDL HPC Profile Application
Extension \cite{jsdl-hpc}:

@@

\begin{lyxlist}{00.00.0000}
\item [{Filename}] - name of the data file on the executing node.
\item [{\emph{Source}}] - source where the file will be taken from before
execution.
\item [{\emph{Target}}] \emph{-} destination the file will be delivered
to after execution.
\end{lyxlist}
\item [{\emph{Input}\ (POSIX,HPC)}] - file to be used as standard input
for the executable.
\item [{\emph{Output}\ (POSIX,HPC)}] - file to be used as standard output
for the executable.
\item [{\emph{Error}\ (POSIX,HPC)}] - file to be used as standard error
for the executable.
\item [{\emph{MemoryLimit}\ (POSIX)}] - amount of physical memory needed
for execution.
\item [{\emph{TotalPhysicalMemory}}] - same as MemoryLimit.
\item [{\emph{IndividualPhysicalMemory}}] - same as MemoryLimit.
\item [{\emph{CPUTimeLimit}\ (POSIX)}] - maximal amount of CPU time needed
for execution.
\item [{\emph{TotalCPUTime}}] - same as CPUTimeLimit.
\item [{\emph{IndividualCPUTime}}] - same as CPUTimeLimit.
\item [{\emph{WallTimeLimit}\ (POSIX)}] - amount of clock time needed
for execution.
\item [{\emph{TotalCPUCount}}] \emph{-} number of CPUs needed for execution.
\item [{\emph{IndividualCPUCount}}] - same as 6.
\item [{\emph{Environment}\ (POSIX)}] - environment variable.
\end{lyxlist}

@@

\begin{itemize}
\item \textit{hardsize} -size of file system for storing cached data
\item \textit{hardfree} - amount of disc space available on that file system
\item \textit{softsize} - if cache exceeds this size files are started being
removed 
\item \textit{softfree} - space left till softsize (can be negative)
\item \textit{claimed} - space used by files claimed by running jobs
\item \textit{unclaimed} - space used by files not being currently used
by any job
\end{itemize}
\item \textit{\#\#\#\#\#\#\#\#.info} - stores state of file (\textit{\#\#\#\#\#\#\#\#}
stands for 8 digits). State is represented by one character:\\
\textit{c} - just created, content is empty.\\
\textit{f} - failed to download (treated same as 'c').\\
\textit{r} - ready to be used, content is valid.\\
\textit{d} - being downloaded. 'd' is followed by identifier of application/job
downloading that file. During content's download this file has write
lock set.
\item \textit{\#\#\#\#\#\#\#\#.claim} - stores list of identifiers of applications/jobs
using this file. Identifiers are stored one per line.
\item \textit{\#\#\#\#\#\#\#\#} - files storing content of corresponding
URL. These can be stored in separate directory.
\end{itemize}
Files \textit{list, old , new} and \textit{\#\#\#\#\#\#\#\#.info}
has to be stored on file system which has support for files' locking.

@@

If a job requests an input file which is subject for caching, it is
stored in the cache directory instead of the SD. The file is made
available to the job by either soft-linking it in the SD or copying
it to the SD. The latter option is more secure and hence advised.

@@

Before giving access to an already cached file, the A-REX contacts
the original file source to check if the user has sufficient access
rights to the file. Not all protocols make this check possible. Also,
file creation or validity times are checked to make sure the cached
file is fresh. If it is impossible to obtain creation and invalidation
times, the file is invalidated 24 hours after it was downloaded.

@@

\begin{itemize}
\item \textit{libarex.so} - The main module providing main functionality
and web interface. It is implemented as HTTP and SOAP service inside
HED. It is responsible for processing jobs, moving them through states
and running other modules.
\item \textit{downloader} - This is a module responsible for gathering input
files in the SD. It processes the \textit{job.ID.input} file and updates
it.
\item \textit{uploader} - This module is responsible for delivering output
files to the specified SEs and registration at an Indexing Service
(like RLS) as needed. It processes and updates the \textit{job.ID.output}
file.
\item \emph{cache-register} - Utility to register cached data into an Indexing
Service. It reads and modifies cache informational files \emph{old}
and \emph{new} (as described in section \ref{sec:cache}). Configuration
is read directly from the A-REX's configuration file (see section
\ref{SubSection: ConfigFile}). It is run by the A-REX every 5 minutes.
\item \emph{frontend-info-collector} - Utility to gather information about
the frontend. It puts collected information into the \emph{job.ID.diag}
file.
\item \emph{gm-kick} - Sends a signal to the A-REX though a FIFO file to
wake it up. It's used to increase responsiveness of A-REX.
\item \emph{cluster+qju.pl} - Collects and generates information about computing
resource as XML document in NorduGrid and Glue 2 format.
\end{itemize}
The following modules are always run under the Unix account to which
a Grid user is mapped.

\begin{itemize}
\item \textit{smtp-send.sh} and \textit{smtp-send} - These are the modules
responsible for sending e-mail notifications to the user. The format
of the mail messages can be easily changed by editing the simple shell
script \textit{smtp-send.sh}. 
\item \textit{submit-{*}-job} - Here {*} stands for the name of the LRMS.
Currently supported LRMS are PBS/Torque, Condor and SGE. Also \emph{fork}
pseudo-LRMS is supported for testing purposes. This module is responsible
for job submission to the LRMS.
\item \textit{cancel-{*}-job} - This script is for canceling jobs which
have been already submitted to the LRMS.
\item \textit{scan-{*}-job} -This shell script is responsible for notifying
the A-REX about completion of jobs. It's implementation for PBS uses
server logs to extract information about jobs. If logs are not available
it uses the less reliable \emph{qstat} command for that. Other backends
use different techniques.
\end{itemize}
In addition, there is also an administration utility:

@@

Due to historical reasons configuration of the A-REX is split into
2 parts. The small part is located inside HED configuration (see appendix
B) for schema and description of supported elements. It refers to
file containing main part of configuration. The default location of
main configuration file is \textit{/etc/arc.conf .}

@@

requires installation of A-REX and all related software to be accessible

@@

\textit{max\_transfered\_files} - maximal number of files being transferred

@@

are registered. Registration is done to same service used for obtaining
information about file. For this operation credentials of the A-REX

@@

\textit{log} - write information about result into log file and continue

@@

\textit{fail} - write information about result into log file and cancel

@@

a user to which Grid identity is mapped in order to access file system

@@

max\_inactivity} - specifies how long/slow data transfer is allowed

@@

data is received for longer than \textit{max\_inactivity} seconds.

@@

Please read \emph{User Manual {[}??]} for usage instructions.
